### 2019-11-01 thesis/txt/2-models.tex
incorporate kw 44 structure feedback in thesis, some plot advancements

\chapter{Information Retrieval Techniques for Build Logs}
\plan{whole chapter: some reordering and reviewing to do, write about 3 pages about ir technique model classes and ir tasks and other techniques}
\plan{write, 1/3 page, blocked by: rest of chapter}
\label{sec:ir-techniques}
%\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information retrieval techniques ... this section describes the conceptional / theoretical results of our thesis ...
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly.
Additionally, we present a complementary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs, we share our notion of retrievable information within a build log.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/build-log-information.pdf}
	\caption{Information retrievable from build logs}
	\label{fig:build-log-information}
\end{figure}
\section{Retrievable Information in Build Logs}
\label{sec:bli}
Continuous integration build logs contain a huge amount of information about the stages in the CI build they correspond to.
In this section, we present a model, which more precisely describes our notion of a information retrievable from a CI build log.
Figure \ref{fig:build-log-information} shows the model.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/ir-technique.pdf}
	\caption{Model for an Information Retrieval Technique}
	\label{fig:model-ie-technique}
\end{figure}
\section{Characteristics of Information Retrieval Techniques}
\label{sec:blirt}
\plan{describe model classes, 1 page text}
For this thesis, we want to evaluate different information retrieval techniques when applied to CI build logs.
We focus on techniques that do not require to parse the structure of a whole build log, but rather focus on extracting just one specified information.
We refer to such a technique as BuildLogInformationRetrievalTechnique (BLIRT).

@@

\section{Information Retrieval Tasks}
\label{sec:blir-tasks}
\plan{what do we need here?, come up with model if we keep this, 1 page text max.}
Previous sections gave various examples of situations where a developer or a researcher wants to retrieve information from a build log.
The goal of this work is to clarify which BLIRT is suited for their specific task.
In this section we structure we present different important characteristics of such an IR task.
You can see the model in Figure \ref{fig:model-ir-task}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/ir-task.pdf}
	\caption{Model for an Information Retrieval Task}
	\label{fig:model-ir-task}
\end{figure}

A InformationRetrievalTask
\begin{itemize}
	\item has a specific amount and kind of \textbf{Configuration} available.
	      This represents how much data is available for the technique to learn what to extract.
	      For a BLIRT to be suited the available configuration has to match the needed configuration of the technique.
	\item is connected to a \textbf{TextProcessor} retrieving the output.
	      This can be for example a developer reading the output or a program.
	      Different TextProcessors require different accuracies in the output, e.g. a program needs exact retrievals to process automatically, while a developer can cope easier with falsely extracted lines.
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

@@

\section{Regular Expression Program Synthesis by Example}
\label{sec:expl-pbe}
\plan{2 caro-reviews before finalized}

\subsection{Introduction}
\emph{Programming by Example (PBE)} is a technique where programs are synthesized according to in and output examples provided by the user.
It enables users to create programs without requiring programming knowledge~\cite{mayer2015user}.
Applied to the task of text extraction through regular expressions it relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
We chose to investigate the applicability of generating regular expression programs by example to extracting information from build logs.
In this work, we refer to our interpretation of this technique as \emph{PBE}.
In the following we will explain the concept of how we apply PBE to information retrieval from CI build logs, the concrete implementation is described in \ref{sec:impl-pbe}.
Figure \ref{fig:prose-explanation} shows the instantiation of the BLIRT model for PBE.

\subsection{Method}
In/Output examples (I/O examples) are the main driver of Programming by Example.
When extracting information from build logs the \emph{input} is always a whole build log, i.e. the whole text of the build log file without any preprocessing.
In the implementation, we had to apply a few text replacements as described in \ref{sec:impl-preprocessing}.
The \emph{output} is always a substring of the log file text, representing the substring that should be extracted by the synthesized program when given the corresponding input file.
One or multiple I/O examples are needed as configuration for PBE, they are used to define substring of a log should be extracted.
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm.
Both are described in Section \ref{sec:rw-prose}.
%For execution PBE takes a build file as input and returns a substring of the build file content.

\subsection{Perspective: Usability vs. Technical Aspects}
Generally, we are interested in whether Programming by Example could replace manually writing regular expressions.
Writing and especially maintaining regular expressions is known to be a difficult, tedious and error-prone task~\cite{michael2019regexes}.
There are two ways to approach this question: from the usability side or the technical side.

When studying the applicability of PBE from the \emph{usability} perspective you run into questions like
\begin{itemize}
	\item Are users faster when using PBE compared to manual regex construction?
	\item Are users more comfortable using PBE over manual regex construction?
	\item Are users confident in the resulting regular expression program from PBE?
	\item Are the resulting extractions more or less accurate when using PBE over manual regex construction?
\end{itemize}
Some of these were already evaluated in existing works about PBE and user interaction~\cite{mayer2015user,lau2009why-programming-by-demonstration,miller2001outlier}.
Adequately answering those questions is however highly dependent on the user interface and the corresponding interaction model used during a study.
Developers who are accustomed to crunching on regular expressions in their favorite IDE or Editor might have an adversely negative view on synthesizing programs from examples in an unknown interface.
Mayer et al.~\cite{mayer2015user} presented two system-user interaction models to improve the confidence of users into the programs synthesized from their examples.
We feel such questions of usability can be answered separately from the application domain of build logs.

For our work, we chose to look at the applicability of PBE from the \emph{technical} perspective.
We investigate whether current program synthesis techniques can learn programs for information retrieval tasks in the domain of CI build logs.
Furthermore, we gather insights into how PBE has to be applied to yield useable results and perform best.
Our implementation of PBE is mainly based on the Flash Extract DSL~\cite{le2014flashextract:} of the PROSE library by Microsoft~\cite{prose2019webpage}.
Due to that our technique of text extraction through regular expression programs synthesized is highly influenced by the capabilities of Flash Extract.

\section{Text Similarity}
\label{sec:expl-ts}
\plan{2 caro-reviews before finalized}

\subsection{Introduction}
Text Similarity approaches are more and more used to filter unstructured textual software artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
One common and simple technique the Vector Space Model~\cite{schutze2008introduction}.
We investigate when it is a suitable technique to retrieve information from build logs.
In the following we will explain the concept of how we apply text similarity to information retrieval from CI build logs (TS), the concrete implementation is described in \ref{sec:impl-ts}.
Figure \ref{fig:text-similarity-explanation} shows the instantiation of the BLIRT model for TS.

\subsection{Method}
To configure a retrieval though text similarity we chose to use the same concept of I/O examples as for PBE.
The lines of the output strings of a given example set define our search query.

@@

\section{Simple Keyword Search}
\label{sec:expl-skws}
\plan{2 caro-reviews before finalized}

\subsection{Introduction}
When we are looking for a specific piece of information within a great amount of unstructured information, a first ad-hoc approach is to search for keywords in the text we assume to be near the information we are looking for.
This was one of the most common approaches we took when searching for the build failure reason within a log when creating our data set.
As this is a technique readily available in many tools developers use to view build logs, we want to study when such a simple search for keywords is recommendable for retrieving information from CI build logs.
In the following we will explain how we use simple keyword search to retrieve information from CI build logs (SKWS), the concrete implementation is described in \ref{sec:impl-skws}.
Figure \ref{fig:keyword-search-explanation} shows the instantiation of the BLIRT model for SKWS.

\subsection{Method}
SKWS is configured by giving a set of keywords.
For a retrieval run, we take a whole build log file as input and search for all exact occurrences of the keywords.
As keywords are often not directly describing the desired information but rather adjacent to the desired information, this technique not only retrieves the line with the keyword but also the context around it.
We chose a context of 10 lines above and below the keyword appearance.

\section{Further Techniques}
\label{sec:expl-rlr}
\plan{decide: what to mention here, max 1 page text}

>>>>

\chapter{Chunk Retrieval Techniques for Build Logs}
\plan{whole chapter: some reordering and reviewing to do, write about 3 pages about ir technique model classes and ir tasks and other techniques}
\plan{write, 1/3 page, blocked by: rest of chapter}
\label{sec:ir-techniques}
%\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information retrieval techniques ... this section describes the conceptional / theoretical results of our thesis ...
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly.
Additionally, we present a complementary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs, we share our notion of retrievable information within a build log.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/build-log-information.pdf}
	\caption{Information retrievable from build logs}
	\label{fig:build-log-information}
\end{figure}
\section{Information Chunks in Build Logs}
\label{sec:bli}
Continuous integration build logs contain a huge amount of information about the stages in the CI build they correspond to.
In this section, we present a model, which more precisely describes our notion of a information retrievable from a CI build log.
Figure \ref{fig:build-log-information} shows the model.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\textwidth, clip]{img/ir-technique.pdf}
	\caption{Model for an Information Retrieval Technique}
	\label{fig:model-ie-technique}
\end{figure}
\section{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
\plan{describe model classes, 1 page text}
For this thesis, we want to evaluate different information retrieval techniques when applied to CI build logs.
We focus on techniques that do not require to parse the structure of a whole build log, but rather focus on extracting just one specified information.
We refer to such a technique as BuildLogInformationRetrievalTechnique (BLIRT).

@@

% \section{Information Retrieval Tasks}
% \label{sec:blir-tasks}
% \plan{what do we need here?, come up with model if we keep this, 1 page text max.}
% Previous sections gave various examples of situations where a developer or a researcher wants to retrieve information from a build log.
% The goal of this work is to clarify which BLIRT is suited for their specific task.
% In this section we structure we present different important characteristics of such an IR task.
% You can see the model in Figure \ref{fig:model-ir-task}

% \begin{figure}[htbp]
% 	\centering
% 	\includegraphics[width=\textwidth, clip]{img/ir-task.pdf}
% 	\caption{Model for an Information Retrieval Task}
% 	\label{fig:model-ir-task}
% \end{figure}

% A InformationRetrievalTask
% \begin{itemize}
% 	\item has a specific amount and kind of \textbf{Configuration} available.
% 	      This represents how much data is available for the technique to learn what to extract.
% 	      For a BLIRT to be suited the available configuration has to match the needed configuration of the technique.
% 	\item is connected to a \textbf{TextProcessor} retrieving the output.
% 	      This can be for example a developer reading the output or a program.
% 	      Different TextProcessors require different accuracies in the output, e.g. a program needs exact retrievals to process automatically, while a developer can cope easier with falsely extracted lines.
% \end{itemize}
% \todo{give 2-3 concrete example instantiations here}

@@

\subsection{Regular Expression Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
\plan{2 caro-reviews before finalized}

\emph{Programming by Example (PBE)} is a technique where programs are synthesized according to in and output examples provided by the user.
It enables users to create programs without requiring programming knowledge~\cite{mayer2015user}.
Applied to the task of text extraction through regular expressions it relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
We chose to investigate the applicability of generating regular expression programs by example to extracting information from build logs.
In this work, we refer to our interpretation of this technique as \emph{PBE}.
In the following we will explain the concept of how we apply PBE to information retrieval from CI build logs, the concrete implementation is described in \ref{sec:impl-pbe}.
Figure \ref{fig:prose-explanation} shows the instantiation of the BLIRT model for PBE.

\paragraph{Method}
In/Output examples (I/O examples) are the main driver of Programming by Example.
When extracting information from build logs the \emph{input} is always a whole build log, i.e. the whole text of the build log file without any preprocessing.
In the implementation, we had to apply a few text replacements as described in \ref{sec:impl-preprocessing}.
The \emph{output} is always a substring of the log file text, representing the substring that should be extracted by the synthesized program when given the corresponding input file.
One or multiple I/O examples are needed as configuration for PBE, they are used to define substring of a log should be extracted.
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm.
Both are described in Section \ref{sec:rw-prose}.
%For execution PBE takes a build file as input and returns a substring of the build file content.

% \subsection{Perspective: Usability vs. Technical Aspects}
% Generally, we are interested in whether Programming by Example could replace manually writing regular expressions.
% Writing and especially maintaining regular expressions is known to be a difficult, tedious and error-prone task~\cite{michael2019regexes}.
% There are two ways to approach this question: from the usability side or the technical side.

% When studying the applicability of PBE from the \emph{usability} perspective you run into questions like
% \begin{itemize}
% 	\item Are users faster when using PBE compared to manual regex construction?
% 	\item Are users more comfortable using PBE over manual regex construction?
% 	\item Are users confident in the resulting regular expression program from PBE?
% 	\item Are the resulting extractions more or less accurate when using PBE over manual regex construction?
% \end{itemize}
% Some of these were already evaluated in existing works about PBE and user interaction~\cite{mayer2015user,lau2009why-programming-by-demonstration,miller2001outlier}.
% Adequately answering those questions is however highly dependent on the user interface and the corresponding interaction model used during a study.
% Developers who are accustomed to crunching on regular expressions in their favorite IDE or Editor might have an adversely negative view on synthesizing programs from examples in an unknown interface.
% Mayer et al.~\cite{mayer2015user} presented two system-user interaction models to improve the confidence of users into the programs synthesized from their examples.
% We feel such questions of usability can be answered separately from the application domain of build logs.

% For our work, we chose to look at the applicability of PBE from the \emph{technical} perspective.
% We investigate whether current program synthesis techniques can learn programs for information retrieval tasks in the domain of CI build logs.
% Furthermore, we gather insights into how PBE has to be applied to yield useable results and perform best.
% Our implementation of PBE is mainly based on the Flash Extract DSL~\cite{le2014flashextract:} of the PROSE library by Microsoft~\cite{prose2019webpage}.
% Due to that our technique of text extraction through regular expression programs synthesized is highly influenced by the capabilities of Flash Extract.

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
\plan{2 caro-reviews before finalized}

Text Similarity approaches are more and more used to filter unstructured textual software artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
One common and simple technique the Vector Space Model~\cite{schutze2008introduction}.
We investigate when it is a suitable technique to retrieve information from build logs.
In the following we will explain the concept of how we apply text similarity to information retrieval from CI build logs (TS), the concrete implementation is described in \ref{sec:impl-ts}.
Figure \ref{fig:text-similarity-explanation} shows the instantiation of the BLIRT model for TS.

\paragraph{Method}
To configure a retrieval though text similarity we chose to use the same concept of I/O examples as for PBE.
The lines of the output strings of a given example set define our search query.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
\plan{2 caro-reviews before finalized}

When we are looking for a specific piece of information within a great amount of unstructured information, a first ad-hoc approach is to search for keywords in the text we assume to be near the information we are looking for.
This was one of the most common approaches we took when searching for the build failure reason within a log when creating our data set.
As this is a technique readily available in many tools developers use to view build logs, we want to study when such a simple search for keywords is recommendable for retrieving information from CI build logs.
In the following we will explain how we use simple keyword search to retrieve information from CI build logs (SKWS), the concrete implementation is described in \ref{sec:impl-skws}.
Figure \ref{fig:keyword-search-explanation} shows the instantiation of the BLIRT model for SKWS.

\paragraph{Method}
SKWS is configured by giving a set of keywords.
For a retrieval run, we take a whole build log file as input and search for all exact occurrences of the keywords.
As keywords are often not directly describing the desired information but rather adjacent to the desired information, this technique not only retrieves the line with the keyword but also the context around it.
We chose a context of 10 lines above and below the keyword appearance.

\subsection{Random Line Retrieval (RLR)}

\subsection{Other Techniques}
\label{sec:expl-rlr}
\plan{decide: what to mention here, max 1 page text}

### 2019-11-01 thesis/txt/3-data-set.tex
incorporate kw 44 structure feedback in thesis, some plot advancements

\chapter{The Failing Build Logs Data Set}
\includegraphics[page=5, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}

@@

\section{Log Collector}
To collect a broad range of build logs we built the \texttt{Log Collector} using ruby.

\subsection{Sampling Repositories}
Our first task was to determine a set of repositories to query logs from.
Our \texttt{GHTorrentParser} can query the GHTorrent~\cite{gousios2013ghtorrent} data set for the most popular languages on github and the most popular repositories for a given language.
\emph{Popularity} in this case is defined as the number of watches.
Our \texttt{TravisRequester} can then check for a given repository whether it uses Travis CI.

\subsection{Sampling Builds}
The \texttt{LogCollector} uses the \texttt{TravisRequester}, our tool querying the Travis API~\cite{travisci2019apidoc}, to obtain the newest builds for a given repository.
As most builds on Travis CI are successful \mention{citation needed} we use a stratified sampling approach: \texttt{TravisRequester} saves the found builds in logs in buckets according to their state.
We encountered the following states during our data collection passes:
\begin{itemize}
	\item \todo{fill /5 mins/}
\end{itemize}
\texttt{TravisRequester} saves up to a configured number of builds per state and searches through a configured number of maximum checked builds.

\subsection{Sampling Logs}
For each build the \texttt{TravisRequester} then selects a log to download.
Logs in Travis CI are not directly attributed to \texttt{build}s, but to \texttt{job}s.
A single build can consist of multiple jobs, e.g.\ building and executing tests in various different testing environments.
\texttt{TravisRequester} therefore queries each build for a job, which has the same state.
A failed build can have successful job executions, as just one failed job leads to the whole build being marked as failed.
For the selected jobs, our tool then queries the Travis API V3 and obtains the respective build log.

@@

\section{Collection Process}
For our initial data collection to get an impression of build logs from various projects and languages we used the \texttt{LogCollector} to gather from the 30 most popular languages, up to 3 repositories using Travis CI each and 3 logs per state for each of those repositories.
For the \emph{Failing Build Log Data Set} we again collected from the same selection of repositories though saved only 10 logs of the state \emph{failed} for each repository.
our impression?
\todo{image!}

\section{Build Failure Reason}
\todo{WHY? one common information developers and researches might want to extract, intentionally not focus on one structurally clearly defined retrieval   cause that would be simple for PROSE retrieval -> better comparability}

\section{Data Model}
For our study, described in Chapter \ref{sec:study}, we want to investigate the retrieval capabilities of three techniques. Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}. Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output. Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples.

\subsection{Example Set}
Our data set is organized in example sets. One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Each example set has an identifying \emph{save name}, a \emph{build log information} that is desired to be extracted by this configuration, and a list of \emph{in / output examples}.

\subsection{In / Output Example}
In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
In our data set they always consist of an \emph{input path}, linking to the build log file which is the input for this example.
The \emph{output} is a substring of that build log, the textual representation of the targeted build log information within the input build log.

\subsection{Keywords}
The keyword search is configured by a list of keywords which are searched in the document with plain text search.
In order to compare it to the two other techniques, each in/output example is associated with a list of 1 - 3 keywords that can be found close to the textual representation of the targeted build log information.

\paragraph{Description}
\todo{for simple keyword search - imitate what users would search for ad-hoc}

\todo{learning steps combining keywords -> similar to how devs would learn from reading those models}

\subsection{Categories}
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.
If two retrievals are at the same structural location with the build log, meaning they are surrounded by a similar markings, the fall into the same category.
For most cases, two build failure reason examples which fall into one category are outputted either within the same build step or by the same build step tool.

@@

\subsection{Sending Mails to Developers}
\subsubsection{Introduction}
For the \emph{Failing Build Log Data Set} we analyzed around 800 build logs from different repositories and tried to extract the part of the log which describes why the respective build failed.
As we are not involved in the development of any of the open source projects within our data set we could only rely on our previous experience with different build logs and systems.
We only looked into the logs and did not check the related configurations, so it is well possible that we extracted parts that do describe errors but the respective step failing is ignored by the configuration and the build failed for another reason.

>>>>

\chapter{BUILDS - BUIld Log Data Set}
\includegraphics[page=5, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}

@@

\section{Log Collection}
To collect a broad range of build logs we built the \texttt{Log Collector} using ruby.

\subsection{Repository Sampling}
Our first task was to determine a set of repositories to query logs from.
Our \texttt{GHTorrentParser} can query the GHTorrent~\cite{gousios2013ghtorrent} data set for the most popular languages on github and the most popular repositories for a given language.
\emph{Popularity} in this case is defined as the number of watches.
Our \texttt{TravisRequester} can then check for a given repository whether it uses Travis CI.

\subsection{Build Sampling}
The \texttt{LogCollector} uses the \texttt{TravisRequester}, our tool querying the Travis API~\cite{travisci2019apidoc}, to obtain the newest builds for a given repository.
As most builds on Travis CI are successful \mention{citation needed} we use a stratified sampling approach: \texttt{TravisRequester} saves the found builds in logs in buckets according to their state.
We encountered the following states during our data collection passes:
\begin{itemize}
	\item \todo{fill /5 mins/}
\end{itemize}
\texttt{TravisRequester} saves up to a configured number of builds per state and searches through a configured number of maximum checked builds.

\subsection{Log Sampling}
For each build the \texttt{TravisRequester} then selects a log to download.
Logs in Travis CI are not directly attributed to \texttt{build}s, but to \texttt{job}s.
A single build can consist of multiple jobs, e.g.\ building and executing tests in various different testing environments.
\texttt{TravisRequester} therefore queries each build for a job, which has the same state.
A failed build can have successful job executions, as just one failed job leads to the whole build being marked as failed.
For the selected jobs, our tool then queries the Travis API V3 and obtains the respective build log.

@@

% \section{Collection Process}
% For our initial data collection to get an impression of build logs from various projects and languages we used the \texttt{LogCollector} to gather from the 30 most popular languages, up to 3 repositories using Travis CI each and 3 logs per state for each of those repositories.
% For the \emph{Failing Build Log Data Set} we again collected from the same selection of repositories though saved only 10 logs of the state \emph{failed} for each repository.
% our impression?
% \todo{image!}

% TODO describe somewhere else
% \section{Build Failure Reason}
% \todo{WHY? one common information developers and researches might want to extract, intentionally not focus on one structurally clearly defined retrieval   cause that would be simple for PROSE retrieval -> better comparability}

\section{Data Structure}
For our study, described in Chapter \ref{sec:study}, we want to investigate the retrieval capabilities of three techniques. Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}. Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output. Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples.

\paragraph{Example Set}
Our data set is organized in example sets. One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Each example set has an identifying \emph{save name}, a \emph{build log information} that is desired to be extracted by this configuration, and a list of \emph{in / output examples}.

\paragraph{In / Output Example}
In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
In our data set they always consist of an \emph{input path}, linking to the build log file which is the input for this example.
The \emph{output} is a substring of that build log, the textual representation of the targeted build log information within the input build log.

\paragraph{Keywords}
The keyword search is configured by a list of keywords which are searched in the document with plain text search.
In order to compare it to the two other techniques, each in/output example is associated with a list of 1 - 3 keywords that can be found close to the textual representation of the targeted build log information.

% \paragraph{Description}
% \todo{for simple keyword search - imitate what users would search for ad-hoc}

% \todo{learning steps combining keywords -> similar to how devs would learn from reading those models}

\paragraph{Categories}
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.
If two retrievals are at the same structural location with the build log, meaning they are surrounded by a similar markings, the fall into the same category.
For most cases, two build failure reason examples which fall into one category are outputted either within the same build step or by the same build step tool.

@@

\subsection{Developer Survey}
\subsubsection{Introduction}
For the \emph{Failing Build Log Data Set} we analyzed around 800 build logs from different repositories and tried to extract the part of the log which describes why the respective build failed.
As we are not involved in the development of any of the open source projects within our data set we could only rely on our previous experience with different build logs and systems.
We only looked into the logs and did not check the related configurations, so it is well possible that we extracted parts that do describe errors but the respective step failing is ignored by the configuration and the build failed for another reason.

### 2019-11-01 thesis/txt/5-study.tex
incorporate kw 44 structure feedback in thesis, some plot advancements

\chapter{Technique Comparison Study}
\label{sec:study}
\section{Introduction}
To investigate how different factors \todo{do we list these somewhere?} influence BLIRTs we evaluate the three techniques implemented in our tool on the \emph{Failing Build Log Data Set}.

@@

\section{Method}
\bp{We compare three different techniques: PBE, TS, SKWS (itemize and link to their explanation?) and as a baseline: random line rerieval (RLR): retrieves random lines from the build log, (describe in chapper about BLIRTs??), number of lines: average of output lines from past}
We configure the technique for every example set with an increasing count of I/O examples and then run the retrieval on the next I/O example, the \emph{test I/O example}.
The output from the test I/O example is the oracle for our evaluation.
The examples are sorted and selected chronologically, i.e. the technique is configured with examples from the directly preceeding build logs to extract information from the test log.
The extraction quality of a BLIRT run is evaluated on only one build log, as it should be suitable for every coming example.
Especially the one directly succeeding the configuration.
If the information retrieval there is bad, we expect the user to directly change the configuration again or be dissapointed with his selection of technique.

@@

\subsection{Regular Expression Program Synthesis by Example}

@@

\subsection{Text Similarity}

@@

\subsection{Simple Keyword Search}

@@

\subsection{Random Line Retrieval}

>>>>

\chapter{Empirical Comparison Study}
\label{sec:study}
To investigate how different factors \todo{do we list these somewhere?} influence BLIRTs we evaluate the three techniques implemented in our tool on the \emph{Failing Build Log Data Set}.

@@

\section{Study Design}
\bp{We compare three different techniques: PBE, TS, SKWS (itemize and link to their explanation?) and as a baseline: random line rerieval (RLR): retrieves random lines from the build log, (describe in chapper about BLIRTs??), number of lines: average of output lines from past}
We configure the technique for every example set with an increasing count of I/O examples and then run the retrieval on the next I/O example, the \emph{test I/O example}.
The output from the test I/O example is the oracle for our evaluation.
The examples are sorted and selected chronologically, i.e. the technique is configured with examples from the directly preceeding build logs to extract information from the test log.
The extraction quality of a BLIRT run is evaluated on only one build log, as it should be suitable for every coming example.
Especially the one directly succeeding the configuration.
If the information retrieval there is bad, we expect the user to directly change the configuration again or be dissapointed with his selection of technique.

@@

\subsection{Regular Expression Program Synthesis by Example (PBE)}

@@

\subsection{Common Text Similarity (CTS)}

@@

\subsection{Keyword Search (KWS)}

@@

\subsection{Random Line Retrieval (RLR)}
